\documentclass[11pt]{article}
\usepackage[table,xcdraw]{xcolor}
\usepackage[colorlinks=true]{hyperref}
\hypersetup{colorlinks,urlcolor=blue}
\usepackage{booktabs}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

\newcommand{\hl}{\begin{flushleft}
	\rule{\textwidth}{1}
\end{flushleft}
}


\newcommand{\n}{\vspace{12pt}}
\usepackage{amssymb} 
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{geometry}



\parindent 0in


\begin{document}


\begin{center} ELON UNIVERSITY\\
DEPARTMENT OF ECONOMICS \\
\end{center}

\begin{center}\textbf{{\hspace{-.05in}}
\underline{ECO 111: PRINCIPLES OF ECONOMICS}}
\\ Fall 2016
\\ MW 3:35PM - 5:15PM (111-L)
\\ MW 5:30PM - 7:10PM (111-M)
%\\ McMichael Science Building 104
\end{center}


\textbf{Instructor:} David A. D\'iaz \hspace{4.5cm} \textbf{Email:} \url{ddiaz5@elon.edu}\\
\textbf{Office:}  Koury Business Center 121   \\
\textbf{Office Hours:} Thursday, 2:00PM-3:30PM\\
\textbf{Course Website:} \url{www.elon.edu/moodle} \\
The Moodle web site contains the official course gradebook, announcements, and other supplementary materials.\\

\textbf{Course Objectives:} The purpose of this course is to introduce you to a new way of looking at the world. The course focuses on core economic concepts and introduces you to some basic models that economists use to make sense of what they observe around them. We will also explore how economists analyze the impact of different policies within the context of these models and determine whether the policy will have the intended - or an unintended - outcome. Throughout the course, I hope to increase your interest in economics and the role it can play in your everyday decision making.
\\

\textbf{Recommended Text:} N. Gregory Mankiw, Principles of Economics, $7^{th}$ Edition.


\subsection*{Grading and Course Components}
\begin{itemize}
	\item Participation: 10\%
	\item Quizzes: 20\%
	\item Exam 1 and 2: 20\% each
	\item	Final Exam: 30\% 
\end{itemize}

The grading scale is as follows:
\begin{center}
	\begin{tabular}{ p{3.5cm} p{3.5cm} }
		A: 93 -- 100 &  C+: 77 -- 79.99\\
		A--: 90 -- 92.99 & C: 70 -- 76.99\\
		B+: 87 -- 89.99 & C--: 67 -- 69.99\\
		B: 83 -- 86.99 & D+: 64 -- 66.99\\
		B--: 80 -- 82.99 & D: 60 -- 63.99\\
		& F: $<$ 60
		
		
	\end{tabular}
\end{center}


\textbf{Problem Sets/Quizzes:} There will be six problem sets assigned during the course of the semester, which you are encouraged to work on in groups. These problem sets \textit{will not be collected or graded}, but rather are intended to be good study guides for the exams and will contain a mix of multiple choice and free response questions. However, there \textbf{will be} six graded quizzes containing questions similar to those on the problem sets. The quizzes are to be completed individually, but are open book/note. Quizzes are due at 11:59PM on the dates noted in the course outline below. \textbf{No late submissions will be accepted}, but your lowest quiz grade for the semester will be dropped. \\

\textbf{Exams:} Each exam will cover material presented in class, readings scheduled outside of class, and homework assignments/quizzes. The exam format will be provided a few days prior to a given exam. You should bring a pencil, scantron, and a \textbf{basic} calculator with you to each exam (an example of what constitutes a basic calculator is provided on Moodle -- if in doubt about yours, ask!). Students that arrive more than 30 minutes late will not be permitted to take the exam. 
\\

\textbf{Participation:} During class, we will work on various problems both as a class and individually with your classmates. Your participation grade will be based on your attendance in addition to your effort level when these group problems are posed. Browsing the web, talking about non-class related topics, playing on your phone, talking while the instructor is talking, etc. will negatively impact your grade.
\\

\textbf{Exam Dates:} \\
October 3  \hspace{4.87cm} Exam 1\\
November 2 \hspace{4.55cm} Exam 2\\
December 12 (111-M) \hspace{2.90cm} Final Exam, 5:30PM - 8:30PM \\
December 15 (111-L) \hspace{3cm} Final Exam, 1:00PM - 4:00PM \\



\subsection*{Course Policies}

\textbf{Attendance \& Participation:} Regular attendance and participation is strongly recommended. You are responsible for any notes or class announcements that you may have missed. You should attempt to get these notes/announcements from one of your peers before seeing me. It is expected that you will respect your peers and the instructor with appropriate behavior while in class and that you will arrive to class on time. You should refrain from browsing the web, texting, etc. during class time. \\


\textbf{Missed Exams:} There are no make-up midterm examinations. If you must miss a midterm exam, you may be permitted to transfer the missed credit to the final examination. To qualify for a transfer of credit, you must contact me before the start of the missed midterm examination and provide me with an acceptable explanation. If the reason for your absence could not be foreseen, please make the request as soon as possible thereafter. All requests should be in writing and you may be asked to provide support with suitable documentation. If approved, the weight of the midterm will be placed on the final exam. \\

If you must miss the final exam for any excused reason, it is your responsibility to alert your instructor and Dean as soon as possible to schedule a make-up exam. Please see the University policy regarding final examinations.\\

\textbf{Academic Integrity:} All students are expected to adhere to the \href{http://elon.smartcatalogiq.com/en/2015-2016/Student-Handbook/The-Honor-System-Code-of-Conduct-and-Processes/Section-I-The-Honor-System}{Honor Code}. 
\begin{itemize}
	\item You may use your notes, textbooks, and classmates to complete the problem sets.
	\item No assistance is permitted on quizzes or exams. During exams, however, feel free to ask your instructor for clarification. The use of cell phones, computers, or any other unauthorized device (e.g., an unapproved calculator) during examinations is an explicit violation of the honor code.
\end{itemize}

\begin{table}[h!]
	\centering
\begin{tabular}{@{}|l|l|l|l|@{}}
	\toprule
	\textbf{Date} & \textbf{Topic}                                                                           & \textbf{Readings}  & \textbf{Notes} \\ \midrule
	8/31          & Introduction to Economics                                                                & Ch. 1; Ch. 2              &            \\ \midrule
	9/5          & Trade                                                                                    & Ch. 3              &                \\ \midrule
	9/7          & Supply and Demand                                                                        & Ch. 4          &            \\ \midrule
	9/12          & Market Equilibrium and Efficiency                                                & Ch. 7              &  Quiz 1 Due on 9/16 \\ \midrule
	9/14          & Elasticity                                                                               & Ch. 5              &                \\ \midrule
	9/19          & Price Controls and Taxes                                                   & Ch. 6       &                \\ \midrule
	9/21          & Government Policy and Welfare                                        & Ch. 8       &                \\ \midrule
	9/26          & Externalities; Public Goods                                                           & Ch. 10; Ch. 11     &               \\ \midrule
	9/28          & Exam 1 Review                                                               &         & Quiz 2 Due on 9/28  \\ \midrule
	\rowcolor[HTML]{009901} 
	10/3          & \multicolumn{3}{c|}{\cellcolor[HTML]{009901}Exam 1}                                                                            \\ \midrule
	10/5          & The Costs of Production                                                                 & Ch. 13             &             \\ \midrule
	10/10          & Perfect Competition                                                                      & Ch. 14             &              \\ \midrule
	10/12          & Monopoly                                                                                 & Ch. 15             &       Quiz 3 Due on 10/16         \\ \midrule
	10/19          & \begin{tabular}[c]{@{}l@{}}Monopolistic Competition;\\ Oligopoly\end{tabular}            & Ch. 16; Ch. 17     &                \\ \midrule
	10/24          & Measuring a Nation's Income                                                             &            Ch. 23        & \\ \midrule
	
	10/26          & Measuring the Cost of Living                                                                                   &       Ch. 24             &     Quiz 4 Due on 10/30           \\ \midrule
	10/31          & Exam 2 Review                                                       &  &                \\ \midrule
	\rowcolor[HTML]{009901} 
	11/2          & \multicolumn{3}{c|}{\cellcolor[HTML]{009901}Exam 2}                                                                            \\ \midrule
	11/7           & Economic Growth                                                     & Ch. 25            &                \\ \midrule
	11/9           & The Solow Model                                                                          & Solow Model videos &    \href{https://www.youtube.com/watch?v=gm2d5s3ZWYU}{Video 1}; \href{https://www.youtube.com/watch?v=e7xBy7zvyN8}{Video 2}  \\ \midrule
	11/14           & \begin{tabular}[c]{@{}l@{}}Savings, Investment, \\ and the Financial System\end{tabular} & Ch. 26             &  \\ \midrule
	11/16           & Unemployment   & Ch. 28   &  Quiz 5 Due on 11/20              \\ \midrule
	11/28           & The Monetary System   & Ch. 29   &                \\ \midrule
	11/30           & Money Growth and Inflation                                                               & Ch. 30             &                \\ \midrule
	12/5          & Aggregate Demand and Supply                                                              & Ch. 33             &                \\ \midrule
	12/7          &  Monetary and Fiscal Policy                                            & Ch. 34             &           Quiz 6 Due on 12/7    \\ \bottomrule
\end{tabular}
\end{table}
\end{document}
